### 2019-01-15 paper/sections/randomtesting.tex

\section{Type-Driven Random Generation} \label{sec:randomtesting}

In this section we briefly describe the common technique used for automatically
deriving random data generators in Haskell using a type-driven approach.
%
Then, we introduce its main drawbacks by considering two scenarios where this
technique gives poor results in practice.
%
We remark that, altough this work makes focus on Haskell data types, this
technique is general enough to be applied to most programming languages with
some level of support for composite types.

Haskell is a strongly typed programming language with a powerful type system.
%
It lets programmers encode a considerable amount of information about the
properties of their systems using data types that can be checked at compilation
time.
%
One of its key aspects is the support for Algebraic Data Types (ADTs).
%
Essentially, an ADTs is a composite type defined by combining other types.
%
In the most basic conception, types can be combined by \textbf{sums} (also known
as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
(or tuples) of other data types.
%
To exemplify this, consider the following definition of the data type |Exp|
encoding integer expressions:

\begin{code}
data Exp  =   Val  Int
          pp  Add  Exp  Exp
          pp  Mul  Exp  Exp
\end{code}

In the previous definition, we declare |Exp| as the sum of three possible
classes of values: |Val| represents literal values, whereas |Add| and |Mul|
represent the addition and multiplication of two integer expressions,
respectively.
%
In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} and are
used to distiguish which variant of the data type we are reffering to.
%
Each data constructor is then defined as a product of zero or more types known
as \emph{fields}.
%
In particular, |Val| contains a field of type |Int|, while |Add| and |Mul|
contain two fields of type |Exp| representing the operands of each operation.
%
Note that |Exp| is used as a field of at least one of its data constructors
(case |Add| and |Mul|), making it a recursively defined ADT.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
using |Exp| data constructors as |(Add (Val 2) (Mul (Val 5) (Val 6)))|.
%
Furthermore, an evaluation function from |Exp|s to integer values can be defined
very idiomatically:

\begin{code}
eval :: Exp -> Int
eval (Val n)    = n
eval (Add x y)  = eval x  +  eval y
eval (Mul x y)  = eval x  *  eval y
\end{code}

In the previous definition, |eval| is described using \emph{pattern matching}
over each possible kind of value.
%
For the input case of a literal values, we simply return the value contained in
the |Val| constructor.
%
On the other hand, if the input value matches either an |Add| or a |Mul| data
constructor, where |x| and |y| are \emph{pattern variables} that match any
subexpression, we recursively evaluate these subexpressions, combining them with
the apropiate operation on each case.

\subsection{\textbf{Type-Driven Generation of Random Values}}

In order to perform random testing of a Haskell codebase involving user defined
data types, most approaches require the user to provide a random data generator
for each one of them.
%
This tends to be a cumbersome and error prone task that closely follows the data
type structure.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Exp|:

\begin{code}
genExp :: Gen Exp
genExp = sized gen
  where
    gen 0 = Val <$> arbitrary
    gen n = oneof
      [  Val  <$> arbitrary
      ,  Add  <$> gen pred   <*> gen pred
      ,  Mul  <$> gen pred   <*> gen pred ]
\end{code} %$

This random generator is defined using \quickcheck's combinator |sized :: (Int
-> Gen a) -> Gen a| to paremetrize the generation process up to an external
integer number known as the \emph{generation size}.
%
This parameter is used to limit the maximum amount of recursive call that this
random generator can perform.
%
When called with a positive generation size (case |gen n|), the generator is free
to randomly pick with uniform probability among any data constructor of |Exp|.
%
In the case it picks to generate a |Val| data constructor, it also generates a
random |Int| value using the standard overloaded generator |arbitrary|
(\quickcheck provides standard random generators for most base types like |Int|,
|Bool|, etc.).
%
On the other hand, when it randomly picks to generate either an |Add| or a |Mul|
data constructor, it also generates independently two random subexpressions
corresponding to the data constructor fields by calling itself recursively (|gen
pred|), ensuring to decrease the generation size by a unit on each recursive
invocation.

This procedure keeps calling itself recursively until the generation size
reaches zero (case |gen 0|), where the generator is constrained to pick only
among terminal data constructors, being |Val| the only possible choice in our
particular case.
%
Strictly decreasing the the generation size by one on each recursive call
results in a useful property over the generated data: every generated value has
at most |n| levels, where |n| is the generation size set by the user.
%
This property enables us to model the generation process using the theory of
branching processes introduced by \tocite, and extended in this work as
described in Section \ref{sec:synthesis}.

Having discussed the previous random generator definition, it easy to understand
how to extend this generation procedure to any data type defined in an algebraic
fashion.
%
Fortunately, there exists different meta-programming tools to avoid the user
from having to mechanically write random generators over and over again for each
user-defined ADT.
%
The simplest tool for such purpose is \megadeth.
%
Given the name of the target data type, it synthesizes a random generator for it
that behaves pretty much like the previously defined one.
%
However, picking among different data constructors with uniform probability can
lead to a generation process biased to generate (in average) very small values,
regardless of the generation sized set by the user \tocite.

\dragen is meta-programming tool conceived to mitigate this problem.
%
Instead of setting a uniform generation probability of data constructors, this
tool considers the scenario where each one can be generated following a
different (although fixed over time) generation probability.
%
Then, this tool uses the theory of branching processes to analitically predict
the average distribution of data constructors generated on each random value.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation probability of each data constructor in
order to obtain a particular distribution of values that can be specified by the
user, offering a more flexible testing environment while still being mostly
automated.

Both \megadeth and \dragen synthesize random generators that are theoretically
capable to generate the whole space of values of the target data type, provided
that we set a generation size suficiently big.
%
However, the limitations arise quickly when we consider that the underlying
generation model is essentially the same: they pick a single data constructor
and recusively generate each required subexpression independently.
%
In practice, this procedure is too generic to generate complex data that is
useful enough to be used for random testing.

In this work we identify two sources of complications that appear when using the
aforementioned automatic derivation tools to obtain random data generators:

\begin{enumerate}
\item The target code behaves differently over inputs matching very specific
  patterns.
\item The target code behaves differently according to some dependency over
  subexpressions of its inputs.
\end{enumerate}

To exemplify the first point, suppose we want to use randomly generated |Exp|s
to test a property comprising the following function:

\begin{code}
foo :: Exp -> Exp
foo (Add (Add x (Val 50)) (Add (Val 25) y))
  = error "pattern #1"
foo (Mul (Mul x (Val 50)) (Mul (Val 25) y))
  = error "pattern #2"
foo x = x
\end{code}

This function behaves mostly like an identity function, except for two very
specific patterns of input values defined using three nested levels of pattern
matching.
%
This is, |foo| does not only matches against the root data constructor, but also
against the data constructors of its subexpressions and sub-subexpressions.

If we consider that \quickcheck's default implementation for random |Int|s pick
a number in the interval $[-n, n]$ with uniform distribution, we need to be able
to identify which one is a sensible generation size to set based on wether our
code branches on numbers or not, otherwise our random generator will not be able
to generate random values to match against such numbers, leaving fragments of
code completely untested.
%
Suppose that we pick the a generation size $50$, i.e. the minimum size that is
big enough to produce an |Int| number equal to $50$.
%
Under this considerations, the probability of generating a value matching the
first clause of |foo| (and hence triggering the error) are as follows:

\begin{align*}
  P(match(foo_1))
  \!&&=\; &P(Add) \\
  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
\end{align*}

\todo[inline]{AM: COMPLETE}

Despite the fact that the previous example might look artificial, branching
against specific patterns of the input data is a common task.
%
For instance, the balancing function of a Red-Black tree need to consider a
quite specific combination of color, left- and right subtrees and sub-subtrees
in order to preserve the height invariant.
%
Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
needs to consider very specific sequences of instructions that may be regrouped
in a computationally more efficient way.

We believe that if a certain pattern of values is relevant enough to appear in
the codebase being tested, a practical random testing methodology should be able
to produce values satisfying such pattern in a substantial proportion.

\begin{code}
fortyTwo :: Exp
fortyTwo = Val 42

twice :: Exp -> Exp
twice x = Add x x

square :: Exp -> Exp
square x = Mul x x

neg :: Exp -> Exp
neg x = Mul x (Val (-1))

minus :: Exp -> Exp -> Exp
minus x y = Add x (neg y)
\end{code}

% \begin{code}
% data IExp  =   VAL Int
%            pp  ADD IExp IExp
%            pp  IF BExp IExp IExp

% data BExp  =   TRUE | FALSE
%            pp  LE IExp IExp
%            pp  AND BExp BExp
% \end{code}

% \begin{code}
% foo :: IExp -> IExp
% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
% foo x                                = ...

% bar :: BExp -> BExp
% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
% bar x                                = ...
% \end{code}

% \begin{code}
% twice :: IExp -> IExp
% twice e = ADD e e

% equals :: IExp -> IExp -> BExp
% equals x y = AND (LE x y) (LE y x)

% and3 :: BExp -> BExp -> BExp -> BExp
% and3 x y z = AND x (AND y z)
% \end{code}