### 2019-01-20 paper/sections/randomtesting.tex

\section{Random Data Generation in Haskell} \label{sec:randomtesting}

In this section we briefly introduce the common approach for automatically
deriving random data generators in Haskell using a type-driven approach, along
with its main drawbacks.

Haskell is a strongly typed programming language with a powerful type system.
%
It lets programmers encode a considerable amount of information about the
structure of their systems using data types that can be checked at compilation
time.
%
One of its key aspects is the support for Algebraic Data Types (ADTs).
%
Essentially, an ADTs is a composite type defined by combining other types in
terms of \textbf{sums} (also known as \emph{variant types}) and
\textbf{products} (or tuples) of other data types.
%
To exemplify this, consider the following type definition to encoding |Html|
expressions:
%
\begin{code}
data Html  =   Text  String
           pp  Sing  String
           pp  Join  Html    Html
           pp  Tag   String  Html
\end{code}

In the previous definition, we declare |Html| as the sum of four possible
constructions: |Text| represents plain text values. |Sing| and |Tag| represent
singular and paired Html tags, and |Join| concats two expression one after
another.
%
We only encode a very small subset of the actual Html specification for
illustrative reasons.
%
In Haskell, |Text|, |Sing| and |Join| and |Tag| are known as \emph{data
  constructors} (or constructors for short) and are used to distiguish which
variant of the data type we are constructing.
%
Each data constructor is defined as a product of zero or more types known as
\emph{fields}.
%
For instance, |Text| has a field of type |String|, whereas |Join| two recusive
fields of type |Html|.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this reprentation, the expression ``|html_example|'' can be encoded as:
%
\begin{code}
Tag "html" (Join (Join
  (Text "hello") (Sing "hr")) (Text "bye"))
\end{code}
%
Additionally, we can define a function |render| to serialize |Html| values as
follows:
%
\begin{code}
render :: Html -> String
render (Text t) = t
render (Sing t) = "<" ++ t ++ ">"
render (Join x y) = render x ++ render y
render (Tag t x)
  = "<" ++ t ++ ">" ++ render x ++ "</" ++ t ++ ">"
\end{code}

In the previous definition, |render| is described using \emph{pattern matching}
over each possible kind of value.
%
Using pattern matching we can define functions idiomatically by defining
different function clauses for each input pattern we are interested on.
%
Patterns can be defined to match specific constructors, literal values or
variable subexpressions (like |t|, |x| and |y| in the definition of |render|).
%
They can also be nested in order to match very specific patterns of values.

\subsection*{\textbf{Type-Driven Generation of Random Values}}

In order to generate random value of types involving user defined ADTs, most
approaches require the user to provide a random data generator for each one of
them.
%
This is a cumbersome and error prone task that closely follows the data type
structure.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Html|:
%
\begin{code}
genHtml = sized (\size ->
   if size == 0
   then frequency
     [  (2,  Text  <$> arbitrary)
     ,  (1,  Sing  <$> arbitrary) ]
   else frequency
     [  (2,  Text  <$> arbitrary)
     ,  (1,  Sing  <$> arbitrary)
     ,  (3,  Join  <$> smaller genHtml <*> smaller genHtml)
     ,  (4,  Tag   <$> arbitrary <*> smaller genHtml) ])
\end{code} %$
%
This random generator is defined using \quickcheck's combinator |sized| to
paremetrize the generation process up to an external natural number known as the
\emph{generation size}.
%
This parameter is chosen by the user, and we use to limit the maximum amount of
recursive calls that this random generator can perform.
%
When called with a positive generation size, this generator can pick to generate
among any |Html| data constructor of with a explicitly given frequency that can
be chosen by the user (2, 1, 3 and 4 for |Text|, |Sing|, |Join| and |Tag|
constructors, respectively).
%
When it picks to generate a |Val| or a |Sing| data constructor, it also
generates a random |String| value using the standard overloaded generator
|arbitrary| (\quickcheck provides standard random generators for most base types
like |String|, |Int|, |Bool|, etc.). \footnote{%
  The operators |<$>| and |<*>| are used in Haskell to combine values obtained
  from effectful computations (like calling to a random generator) and they are
  not particularly relevant for the point being made in this work.}
%$
On the other hand, when it picks to generate either an |Join| constructor, it
also generates two independent random subexpression recursively, decreasing the
generation size by a unit on each recursive invocation (|smaller genHtml|).
%
The case of random generation of |Tag| constructors follows analogously.

This random process keeps calling itself recursively until the generation size
reaches zero, where the generator is constrained to pick only among terminal
data constructors, being |Text| and |Sing| the only possible choices in our
particular case.

% Strictly decreasing the the generation size by one on each recursive call
% results in a useful property over the generated data: every generated value has
% at most |n| levels, where |n| is the generation size set by the user.
%
% This property enables us to model the generation process using the theory of
% branching processes introduced by \tocite, and extended in this work as
% described in Section \ref{sec:synthesis}.

The previous definition is rather mechanical, except perhaps for the chosen
generation frequencies.
%
In this light, it easy to extend this procedure to any data type defined in an
algebraic fashion.
%
Fortunately, there exists different meta-programming tools to avoid the user
from having to mechanically write random generators over and over again for each
user-defined ADT.
%
The simplest tool for such purpose is \megadeth \tocite.
%
Given a target data type, it synthesizes a random generator for it that behaves
similarly to the one presented above, where the generation frequencies are
defined to be uniform across constructors.
%
However, picking among different data constructors with uniform frequency can
lead to a generation process biased towards generating (in average) very small
values, regardless of the generation sized set by the user \tocite.

\dragen is meta-programming tool conceived to mitigate this problem.
%
Instead of setting a uniform generation probability of data constructors, this
tool this tool uses the theory of branching processes to modelize and predict
analitically the average distribution of data constructors generated on each
random value.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation frequency of each data constructor in order
to obtain a particular distribution of values that can be specified by the user,
providing this way a more flexible testing environment while still being mostly
automated.

Althogh both \megadeth and \dragen synthesize random generators that are
theoretically capable to generate the whole space of values of the target data
type. the limitations arise quickly when we consider that the underlying
generation model is essentially the same: they pick a single data constructor
and recusively generate each required subexpression independently.
%
In practice, this procedure is often too generic to generate random data with
enough structural complexity to be used for testing purposes.

% In this work we identify two sources of additional structure information which
% are not considered by the aforementioned automatic derivation tools to obtain
% better random data generators:

In this work we identify two patological situations which are not properly
handled by the aforementioned derivation tools:

\begin{enumerate}
\item The target code behaves differently over inputs matching specific patterns
  of nested values.
\item The target code encodes a significant portion of its structure on its
  abstract interface.
\end{enumerate}

In Section \ref{sec:hrep} we show how this structural information can be used to
synthesize richer random generators automatically.
%
We proceed to exemplify the previous points in detail.

\subsection*{\textbf{Presence of Complex Pattern Matchings}}

To exemplify the first problematic scenario, suppose we want to use randomly
generated |Htmls|s to test a property comprising the following function:
%
\begin{code}
simpl :: Html -> Html
simpl (Join (Text t1) (Text t2))
  = Text (t1 ++ t2)
simpl (Join (Join (Text t1) x) y)
  = simpl (Join (Text t1) (simpl (Join x y)))
simpl (Join x y) = Join (simpl x) (simpl y)
simpl (Tag t x) = Tag t (simpl x)
simpl x = x
\end{code}

This function simplifies sequences of |Text| constructors into a single big
|Text| constructor.
%
To do so, it has to pattern match against sequences of |Text| constructors
combined by a |Join| constructor using nested patterns (see |simpl| fist and
second clauses).
%
The remaining clauses are only meant to propagate this simplification within
nested expressions.

Ideally, we would like to test each clause of the function |simpl| approximately
the same amount of time each.
%
However, each data constructor is generated independently when using either
\megadeth or \dragen, thus the probability of generating a value satisfying a
nested pattern decreases multiplicatively with the number of constructors we
pattern against to simultaneously.
%
In our tests, we found that the first two clauses of |simpl| get exercised only
approximately $1.5\%$ of the time when using \megadeth to derive a random
generator for |HTML|.
%
On the other hand, the best result we could achieve with \dragen was only able
to exercise the first and second clauses of |simpl| approximately $3\%$ and
$6\%$ of the time, respectively.
%
With both derivation tools, the most of the time was spent testing the trivial
clauses of our function, in view of they pattern match against simpler patterns
of input values.

% In this light, we found that using a random generator obtained with \megadeth to
% test this function will only exercise the first two clauses of |simpl| about
% only $1.5\%$ of the time each, relegating the rest of the generated test cases
% to the remaining and somewhat less interesting clauses, where, remarkably, half
% of the time is spent exercising the last trivial clause.
% %
% Furthermore, using \dragen to ...

% This function behaves very much like an identity function, with the exception
% that it fails for two very specific patterns of input values defined using
% nested pattern matching.
% %
% This is, |foo| does not only matches against the root data constructor, but also
% against the data constructors of its subexpressions and sub-subexpressions.

% \quickcheck's default implementation for random |Int|s pick a number in the
% interval $[-n, n]$ with uniform distribution.
% %
% Hence, in principle we need to recognize what is a suitable generation size,
% which should be big enough for our generator be able to generate the numbers we
% pattern match against to, otherwise we will not be able to generate any value to
% match against |foo| clauses, leaving fragments of code completely untested.
% %
% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
% is big enough to produce an |Int| number equal to $50$.
% %
% Under this consideration, the probability of generating a value matching the
% first clause of |foo| (and hence triggering the first error) results as follows:
% %
% \begin{align*}
%   &P(match(foo\#1))\\
%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
% \end{align*}

% If we use \megadeth to automatically derive a random generator for |Exp|, we
% obtain a uniform generation probability distribution over constructors, i.e.,
% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
% %
% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
% average, we will need to generate over than two million test cases in order to
% be able to test the first clause of |foo| only once.
% %
% This situation can be somewhat improved if we use \dragen to obtain a random
% generator.
% %
% Using this tool, we can optimize the generation probabilities in order to
% benefit the generation of some data constructors over the rest.
% %
% Considering that the first pattern matching of |foo| involves the data
% constructor |Add| as the only recursive one, we can set an target generation
% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
% the rest of the generated data constructors.
% %
% By doing so, the obtained distribution of values results such that
% $P(match(foo\#1)) \approx 1/300000$.
% %
% Although this certainly improves the probability of generating a matching value,
% this probability is not substantial enough to become practical.

% Additionally, by favoring the generation probabilities towards the |Add|
% constructor, we found that the probability of generating a value matching the
% second clause of |foo| (which does not matches against it) also gets diminished
% into an impractival value.

Altough the previous example might seem rather simple, branching against
specific patterns of the input data is a common task.
%
For instance, balancing a Red-Black tree requires to consider specific
combinations of color, left and right subtrees and sub-subtrees in order to
preserve the height invariant \tocite.
%
Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
needs to consider very specific sequences of instructions that may be regrouped
in a computationally more efficient way, to cite a few \tocite.

\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}

A common choice when implementing a data structure is to transfer the
responsability of preserving its invariants to the functions which manipulates
its values.
%
For this purspose, suppose we extend our |Html| data type with the following
basic combinators:
%
\begin{code}
module M where

  div :: Html -> Html
  div inner = Tag "div" inner

  bold :: Html -> Html
  bold inner = Tag "b" inner

  hr :: Html
  hr = Sing "hr"
\end{code}
%
% In the previous definition, the |Html| data type is defined using a single data
% constructor that contains the textual representation of the Html code it
% represents in plain text.
%
These functions encode additional information about the structure of our |Html|
data type in the form of specific Html tags.

Instead of including a new data constructor for each possible Html tag in out
type definition, we defined a mininal representation and then extended it with
an set of high level combinators.
%
% These combinators functions defined over |HTML| are the ones in charge of
% transforming this plain text representation with the invariant that, given valid
% |Html| parameters, they always return a valid |Html| value.
%
This programming pattern is often called a ``shallow embedding'', and can be
found in a variety of Haskell libraries, being \emph{html} \tocite,
\emph{svg-builder} \tocite some examples of this.

As a consequence of this practice, type-driven derivation techniques often fail
to synthesize useful random generators due to that most of the data type
structure has been encoded into its abstract interface of combinators.
%
In our particular case, the chances of generating a |Tag| value representing a
commonly used Html tag such as |div| are extremely low.

% Additionally, if our test suite contains properties constrained by certain
% preconditions, the lack of domain knowleadge may lead in an impractically high
% discard ratio of randomly generated test cases.
% %%
% For instance, suposse we write a property to verify that |render| always outputs
% a valid |Html| value:

% \begin{code}
% prop_render :: Html -> Bool
% prop_render x = valid x ===> valid (parse (render x))
% \end{code}

% In the previous definition we state that |render| always returns a valid |Html|
% value when it is parsed again (|parse (render x)|), provided that we are given a
% valid |Html| as input (|valid x|).
% %
% In practice, not every string of characters constitutes a valid Html given that
% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).

% %
% While testing this property we rarely satisfy its precondition, in which case
% \quickcheck discards the whole test, retrying with a diferent random input,
% degradating this way the testing performance.

% \begin{code}
% text :: String -> Bool
% text str = Text (addEscapes str)
% \end{code}

% We believe that, if certain patterns of values are relevant enough to appear in
% the codebase being tested, a practical random testing methodology should be able
% to produce values satisfying such patterns in a substantial proportion.

So far we have introduced two testing scenarios where type-driven derivation
approaches are unable to capture all the available structure information from
the user codebase.
%
Fortunately, this information can be automatically exploited and used to
generate interesting and more structured random values.

The next section introduces a reprentation model that let us encode the
structure information presented here into our automatically derived random
generators in a modular and flexible way.