### 2019-01-21 paper/sections/hrep.tex
WIP

\section{Extracting Structure} \label{sec:hrep}

In this section we present a compositional representation of values to express
the random generation of values following the internal structure of their data
types along with the structure present on patterns matchings and abstract
interfaces.

The key idea of this work is to represent different structured constructions of
data in a homogeneous way that we call a ``higher-level representation'' (|HRep|
from now on).
%
Instead of generating each data constructor independently, a random generator
derived from this representation might generate composite structured values on
each random choice it performs.
%
% \footnote{The notion of higher-level comes from that the generation process is
%   entirely determined by the type of the chosen type level representation,
%   instead of by a concrete generator defined at the term level.}
%
For this purpose, we use a series of automatically derived data types, each one
representing an atomic unit of information that can be randomly generated and
then reflected back to the corresponding value of the original data type.
%
Later, the user can compose these atomic representations using the provided type
level combinators in different ways into a ``generation specification'' that
completely determines the generation process behavior.

We wil reuse the previously defined data type |Html| and the functions defined
in the previous section to explain the different concepts involved all across
this section.

\subsection*{\textbf{Representing Data Constructors}}

We begin by introducing the simplest data type representation that we can
extract from our codebase: the representation of single data constructor.
%
Each data constructor can be represented by an automatically derived data type
consisting of a single constructor with the same fields as the original, except
for the recursive ones that are abstracted away.
%
In this light, we represent each constructor of the data type |Html| as follows:

\begin{code}
data HRepText  r = ConText  String
data HRepSing  r = ConSing  String
data HRepJoin  r = ConJoin  r r
data HRepTag   r = ConTag   String r
\end{code}

Note that the previous definitions are type parametric over the type parameter
|r|.
%
This allow us to replace |r| with any concrete data type, obtaining different
possible values on each case.
%
For instance, the value (|ConJoin 10 20|) has type |HRepJoin Int|, while the
value (|ConJoin True False|) has type |HRepJoin Bool|.

In practice, this parametricity let us instatiate |r| with the type of the
chosen generation specification (which might be composed of several |HRep|s),
without having to modify anything in the underlying machinery.

Having the |HRep| of each data constructor, we can define an evaluation relation
($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
representation |f| back to the target data type |t|.
%
Then, we simply need to translate each constructor representation back into its
corresponding one, translating the abstracted fields recursively:
%
\begin{alignat*}{4}
  &\evalrep{|ConText|\ s &&}{Html}
    &&= |Text|\ s \\
  &\evalrep{|ConSing|\ s &&}{Html}
    &&= |Sing|\ s \\
  &\evalrep{|ConJoin|\ x\ y &&}{Html}
    &&= |Join|\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
  &\evalrep{|ConTag|\ s\ x &&}{Html}
    &&= |Tag|\ s\ \evalrep{x}{Html}
\end{alignat*}

The missing piece is to automatically synthesize a random generators for each
constructor representation.
%
For this purpose, it is important to consider that each constructor |HRep| has
its recursive fields abstracted away with a type parameter that will be later
instantiated with the generation specification type.
%
Given that this specification is unknown at the derivation time, we parametrize
each |HRep| generator with a random generator |genr| that is used to generate
random values for each recursive field:

\begin{code}
  genText  genr  = ConText  <$> arbitrary
  genSing  genr  = ConSing  <$> arbitrary
  genJoin  genr  = ConJoin  <$> smaller genr <*> smaller genr
  genTag   genr  = ConTag   <$> arbitrary <*> smaller genr
\end{code}

\subsection*{\textbf{Type Level Combinators}}

The next step is to define a series of type level combinators to enable us
combining the automatically derived |HRep|s in several ways.

In first place, we define a type combinator (|term|) to tag a |HRep| to be
terminal, i.e., a representation that is allowed be generated when the
generation size gets exausted:
%
\begin{code}
data (f_term) a = Term (f a)
\end{code}

Additionally, we define a combinator ($\otimes$) to tag a |HRep| with an
explicit generation frequency $n$:

\begin{code}
data (f otimes n) a = Freq (f a)
\end{code}

The previous combinators only include information relevant to the generation
process, in a sense that neither one adds new structure to the final
representation.
%
In this light, they do not alter the evaluation semantics, and we translate them
back to our target data type by evaluating the inner representation:
%
\begin{alignat*}{4}
  &\evalrep{Term\ x &&: |f_term|    &&}{t} &&= \evalrep{x : f}{t} \\
  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
\end{alignat*}

In the previous equations we explicitly annotate (using a colon) the type of the
evaluated term for clarity.
%
Later, to generate these combinators is enough to wrap a generated value from
the inner representation with the apropriate tag:

\begin{code}
gen_f_term     genr  = Term  <$> genf genr

gen_f_times_n  genr  = Freq  <$> genf genr
\end{code}

Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
|HRep|s into a single one using a sum type to represent a random choice between
them:

\begin{code}
data (f oplus g) a = L (f a) | R (g a)
\end{code}

A composite representation built using $(\oplus)$ is transformed back into the
target data type by pattern matching on the data constructor variant and
evaluating the inner |HRep| accodingly:
%
\begin{alignat*}{3}
  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
\end{alignat*}

Generating a composite |HRep| is slightly more complicated than before, as we
need to perform a random choice based on the generation size and the given
frequencies for each sub-represention:

\begin{code}
  gen_f_plus_g genr  = sized (\size ->
    if size == 0
    then frequency
      [ (freq0  at_f,  L  <$> genf  genr  )
      , (freq0  at_g,  R  <$> geng  genr  ) ]
    else frequency
      [ (freq   at_f,  L  <$> genf  genr  )
      , (freq   at_g,  R  <$> geng  genr  ) ])
\end{code} %$

In the previous definition, we reflect the type level frequencies of the types
|f| and |g| (|freq at_f| and |freq at_g|) into term-level values.
%
This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
%
Then we use these frequencies to generate each inner |HRep| in the apropriate
proportion.
%
When the generation size gets exhausted, we reflect the terminal generation
frequency of each inner |HRep| in the same way as before (|freq0 at_f| and
|freq0 at_g|).
%
This time, however, we default the frequency reflection to $0$ for any inner
that not tagged as terminal, avoiding to generate non-terminal constructions in
the last step.

Later, we can use these combinators to create a type synonym |HRepHtml| that
specifies a generation schema equivalent to the one seen in the concrete random
generator of type |Html| presented in Section \ref{sec:randomtesting}:

\begin{code}
type HRepHtml  =       HRepText_term  otimes 2
               oplus'  HRepSing_term
               oplus'  HRepJoin       otimes 3
               oplus'  HRepTag        otimes 4
\end{code}

However, a value of type |HRepHtml| still has its recursive calls abstracted
away---the type parameter |r| is implicit at the definition of |HRepHtml|.
%
We can think of it as a ``single layer'' of representation.
%
To make it able to represent recursive values we need to define a last type
level combinator to ``tie the knot'':

\begin{code}
  data Fix f = Fix (f (Fix f))
\end{code}

This datatype represents the \emph{fixed point} of a parametric data type |f|,
i.e., a data type where each recursive call gets instantiated with iself.
%
Then, to translate fixed points back to our target data type we simply need
translate the inner representation:
%
\begin{align*}
  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
\end{align*}

% Categorically, this generic transformation of fixed points is called a
% \emph{catamorphism}, where our evaluation relation is known as its
% \emph{algebra} \tocite.

Unlike the other combinators, to define a random generator for the fixed point
of a certain representation |f|, we do not to parametrize the generation of the
recursive fields of |f| over an external generator |genr|.
%
% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
Instead, we replace it with our fixed point generator, calling itself
recursively on any recursive field of |f| that might appear inside:

\begin{code}
genFix_f = Fix <$> genf genFix_f
\end{code} %$

This way we obtain a concrete recursive generator for each representation |f|
that we define.
%
Then, we can define a random generator for our target data type simply by
generating a random value of our chosen representation, and transforming it back
to our target by the means of the evaluation relation:

\begin{code}
  genHtml = do  x <- genFix_HRep_Html
                return eval_x_Html
\end{code} %$

So far we have introduced the machinery required to represent the random
generation of a target data type considering only the structure encoded on its
definition.
%
However, this approach can now be easily extended to encode different sources of
structured information.
%
We proceed to introduce two extensions that help to address the problematic
testing scenarios presented in the previous section.

\subsection*{\textbf{Representing Pattern Matchings}}

We can follow a similar reasoning as before to represent the pattern matching
structure from a given function.
%
Consider the nested pattern matchings of function |simpl| defined in the
previous section:

\begin{code}
simpl (Join (Text t1) (Text t2))   = ...
simpl (Join (Join (Text t1) x) y)  = ...
\end{code}

To represent these patterns, we derive a new data for each one of them, whose
fields represent the pattern variables that occur inside, and abstracting away
every pattern variable of type |Html|.
%
Concretely, we define the following data types:

\begin{code}
data HRepSimpl1  r = PatSimpl1 String String
data HRepSimpl2  r = PatSimpl2 String r r
\end{code}

The first pattern of |simpl| contains two pattern variables (|t1| and |t2|) of
type |String| that are included as fields of |PatSimpl1|.
%
Similarly, the second pattern of |simpl| contains a pattern variable |t1| of
type |String| represented by the first field of |PatSimpl2|, along with two
pattern variables (|x| and |y|) of type |Html| that are abstracted in the second
and third fields of |PatSimpl2|.

To transform these representations back to our target data type, we simply need
to expand them it into the concrete value represented by the original pattern,
evaluating its fields back to |Html| as well:
%
\begin{align*}
  &\evalrep{|PatSimpl1|\ t1\ t2}{Html} \\
  &\quad = |Join (Text eval_t1_Html) (Text eval_t2_Html)|\\
  &\evalrep{|PatSimpl2|\ t1\ x\ y}{Html} = \\
  &\quad = |Join (Join (Text eval_t1_Html)) eval_x_Html) eval_y_Html)|
\end{align*}

This way, any value of type |HRepSimpl1| is guaranteed to satisfy the first
pattern matching of |simpl|---the same property follows for |HRepSimpl2|.

The random generation of pattern |HRep|s is defined in the same way as we did
before for representations of data constructors:

\begin{code}
genSimpl1  genr
  = PatSimpl1  <$> arbitrary  <*> arbitrary
genSimpl2  genr
  = PatSimpl2  <$> arbitrary <*> smaller genr <*> smaller genr
\end{code} %$

Finally, we can join the pattern matching representation of each clause of
|simpl| into a single one:

\begin{code}
type HRepSimpl  = HRepSimpl1 oplus''  HRepSimpl2
\end{code}

\subsection*{\textbf{Representing Abstract Interfaces}}

To introduce the higher level representation of module abstract interfaces,
consider module |M| defininf |Html| combinators introduced in the previous
section.

% \begin{code}
% module M where

% ten :: Exp
% ten = Val 10

% square :: Exp -> Exp
% square x = Mul x x

% minus :: Exp -> Exp -> Exp
% minus x y = Add x (Mul y (Val (-1)))
% \end{code}

We can represent each function of |M| that returns a value of type |Html|
following the same idea as before, deriving a data type with a single data
constructor for each one of them:

\begin{code}
data HRepdiv        r = Fundiv r
data HRepbold       r = Funbold r
data HRephr         r = Funhr
\end{code}

In this case, each single constructor will have as fields the types of the
inputs of the function that they represent.
%
As before, we abstract away any field representing and input of type |Html| with
a type parameter |r|.

To evaluate these representations, we simply use the values on its fields as
input parameter of each original function, returning its result.
%
\begin{alignat*}{3}
  &\evalrep{|Fundiv|\  &&x  &&}{Html} = |div eval_x_html| \\
  &\evalrep{|Funbold|\ &&x &&}{Html} = |bold eval_x_html| \\
  &\evalrep{|Funhr|    &&  &&}{Html} = |hr|
\end{alignat*}

Note that, by doing this, the generation process inherits any patology that the
functions we use to generate values might have.
%
For instance, if the function |div| would happen to be non-terminating for some
inputs, our generation process could suffer from this as well.

Furthermore, the generation procedure for abstract interface representations
follows the same pattern as before:

\begin{code}
  gendiv   genr  = Fundiv <$> smaller genr
  genbold  genr  = Funbold <$> smaller genr
  genhr    genr  = return Funhr
\end{code} %$

And we will also define a type synonym to join all the representations of the
module |M| into a single one.

\begin{code}
type HRepM  = HRepdiv oplus''  HRepbold oplus''  HRephr
\end{code}

Finally, we can put all the derived machinery together into a generation
specification |HtmlS|, assigning (possibly) different generation frequencies to
each individual |HRep| we combine:

\begin{code}
type HtmlS  =       HRepHtml   otimes 4
            oplus'  HRepSimpl  otimes 2
            oplus'  HRepM
\end{code}

This previous definition can be interpreted graphically as it is shown in the
Figure \ref{fig:hrep}, where curly arrows represent the structural information
extracted using meta-programming.

\begin{figure*}[t]
  \centering
  \input{tikz/hrep.tex}
  \caption{Higher-level representation of the data type |Html|, defined using
    structural information from the function |simpl| and the abstract interface
    of the module |M|.}
  \label{fig:hrep}
\end{figure*}

% \begin{figure}[t]
%   \centering
%   \input{tikz/val.tex}
%   \caption{Some caption.}
%   \label{fig:val}
% \end{figure}

We want to remark that, for space reasons, we were only able to introduce the
representation of a rather simple target data type.
%
In practice, this reasoning can be extended to mutually recursive and parametric
types as well.

Overall, this approach offers significant advantages over the usual type-driven
derivation of random generators:
%
\begin{itemize}
\item \textbf{Composability:} we can combine different atomic representations
  using different structure information sources depending on what property or
  sub-system we need to verify using randomly generated values.
  %
\item \textbf{Extensibility:} the developer can derive representations for new
  sources of structure information and combine them with the existing ones
  simply by adding them to the existing generation specification of the target
  data type.
  %
\item \textbf{Predictability:} using branching processes theory, it is possible
  to predict the average distribution of generated values in terms of number of
  constructors.
  %
  This prediction is completely modular, and can be obtained for any composite
  representation obtained using the automatically derived |HRep|s and the
  provided combinators.
\end{itemize}

\todo[author=AM, inline]{Maybe we can show an example random value from the representation
  and its corresponding target value }

% \begin{align*}
%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
% \end{align*}
%
% \begin{code}
% class (rep down target) where
%   step :: rep target -> target
% \end{code}
%
% \begin{code}
% data (Term f) a = TagTerm (f a)

% \begin{code}
% instance (HRep_Val down Exp) where
%   step (Mk_Val n) = Val n

% instance (HRep_Add down Exp) where
%   step (Mk_Add x y) = Add x y

% instance (HRep_Mul down Exp) where
%   step (Mk_Mul x y) = Mul x y
% \end{code}

% \begin{code}
% instance (f otimes n down Exp) where
%   step (Freq f) = step f
% \end{code}

% \begin{code}
% instance Arbitrary1 (f otimes n) where
%   liftArbitrary gen_HRep = Tag <$> gen_HRep
% \end{code} %$

% instance (Term f down Exp) where
%   step (TagTerm f) = step f
% \end{code}

% \begin{code}
% instance (HRep_ten down Exp) where
%   step Mk_ten = ten

% instance (HRep_square down Exp) where
%   step (Mk_square x) = square x

% instance (HRep_minus down Exp) where
%   step (Mk_minus x y) = minus x y
% \end{code}

% \begin{code}
% instance Arbitrary1 HRep_ten where
%   liftArbitrary gen_HRep = pure Mk_ten

% instance Arbitrary1 HRep_square where
%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep

% instance Arbitrary1 HRep_minus where
%   liftArbitrary gen_HRep
%     = Mk_minus <$> gen_HRep <*> gen_HRep
% \end{code} %$

% \begin{code}
% instance (HRep_foo_1 down Exp) where
%   step (Mk_foo_1 x y)
%     = Add (Add x (Val 50)) (Add (Val 25) y)

% instance (HRep_foo_2 down Exp) where
%   step (Mk_foo_2 x y)
%     = Mul (Val 50) (Mul (Val x) y)
% \end{code}

% \begin{code}
% instance Arbitrary1 HRep_foo_1 where
%   liftArbitrary gen_HRep
%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep

% instance Arbitrary1 HRep_foo_2 where
%   liftArbitrary gen_HRep
%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
% \end{code} %$

% \begin{code}
% instance (f oplus g down Exp) where
%   step (L f) = step f
%   step (R g) = step g
% \end{code}

% \begin{code}
% instance Arbitrary1 (f oplus g) where
%   liftArbitrary gen_HRep
%     = frequency
%       [ (freq at_f, L <$> gen_HRep)
%       , (freq at_g, R <$> gen_HRep) ]
% \end{code} %$

% \begin{figure}[b]
%   \centering
%   \input{tikz/val.tex}
%   \caption{Higher level representation of the data type |Exp|, defined using
%     structural information from the function |foo| and the abstract interface of
%     the module |M|.}
%   \label{fig:hrep}
% \end{figure}