### 2019-01-23 paper/abstract.tex
Abstract

\begin{abstract}

  % Status
  Automatic generation of random values described by abstract data types (ADTs)
  is often a hard task.
  % data types (ADTs), % ---specially when considering ADTs which
  % represent HTML pages, images, programs or other structural rich objects.
  % Situation
  State-of-the-art random testing tools often can automatically synthesize
  random data generators based on ADTs definitions.
  %
  In that manner, generated values comply with the structure described by
  ADTs---something that proves useful when testing software which expects
  complex input like web pages, images, or even programs.
  %
  % Problem
  However, it sometimes becomes necessary to generate structural richer ADTs
  values in order to test deeper software layers.
  %
  % Solution
  In this light, we propose to leverage static information found in the
  codebase as a manner to improve the generation process.
  %
  Namely, our generators are capable to consider how programs branch on input
  data as well as how ADTs values are built via interfaces.
  %
  % Concretely, we propose to consider not only ADTs definitions, but also how
  % programs branch on input data as well as how ADTs values get manipulated via
  % interfaces.
  %
  We implement a tool, called {\dragenp}, responsible to synthezise generators
  for ADTs values while providing compile-time guarantees about their
  distributions.
  %
  Our compile-time predictions allow {\dragenp} to provide an heuristic that
  tries to adjust the distribution of generators to what developers might want.
  \todo[inline,author=Ale]{Case study}

\end{abstract}