### 2019-01-29 paper/synthesis.tex
WIP

\section{Predicting Distributions} \label{sec:synthesis}

Characterizing the distribution of values of an arbitrary random generator is a
hard task.
%
It requires to model every random choice that a generator could possibly make to
generate a value.
%
In a recent work \cite{DBLP:conf/haskell/MistaRH18}, we have shown that it is
possible to \emph{analitically} predict the average distribution of data
constructors produced by a random generators automatically derived considering
only ADT definitions---like the one presented on Section
\ref{sec:randomtesting}.
%
For this purpose, we found that random generation of ADT values can be
characterized using the theory of \emph{branching processes} \cite{gw1875}.
%
This probabilistic theory was originally conceived to predict the growth and
extinction of royal family trees the Victorian Era, being later applied to a
wide variety of research areas.
%
In this work, we adapt this model to also predict the average distribution of
values of random generators derived considering structural information coming
from functions pattern matchings and abstract interfaces.

Essentially, a branching process is a special kind of Markov process that models
the evolution of a population of \emph{individuals of different kinds} across
discrete time steps known as \emph{generations}.
%
Each kind of individual is expected to produce an average number of offspring of
(possibly) different kinds from one generation to the next one.
%
Mista el at. \cite{DBLP:conf/haskell/MistaRH18} show that branching processes
can be adapted to predict the generation of ADT values by simply considering
each data constructor as an individual on its own.
%
In fact, any ADT value can be seen as a tree where each node represents a root
data constructor and has its sub-expressions as sub-trees---hence note the
similarity with family trees.
%
In this light, each tree level of a random value can be seen as a generation of
individuals in this model.

We characterize the numbers of constructors that a random generator produces in
the $n$-th generaton as a vector $G_n$, a vector that groups the number of
constructors of each kind produced in that generation---in our |Html| example,
this vector has four components, i.e., one for each constructor.
%
%We note $G_n$ to the vector of real numbers that represents the distribution of
%
From branching processes theory, the following equation captures the expected
distribution of constructors at the generation $n$, noted $E[G_n]$, as follows:
%
\begin{align}
  E[G_n]^T = E[G_0]^T \cdot M^n
  \label{eqn:prediction}
\end{align}

Vector $E[G_0]$ represents the initial distribuition of constructors that our
generator produces, which simply consists of the generation probability of each
one.
%
The interesting aspect of the prediction mechanism is encoded in the matrix $M$,
known as a the \emph{mean matrix} of this stochastic process.
%
$M$ is a squared matrix with as rows and columns as different data constructors
involved in the generation process.
%
Each element $M_{i,j}$ of this matrix encodes the average number of data
constructors of kind $j$ that gets generated in a given level, provided that we
generated a constructor of kind $i$ at the previous level.
%
In this sense, this matrix encodes the ``branching'' behavior of our random
generation from one generation to the next one.
%
Each element of the matrix can be automatically calculated by exploiting ADT
definitions \cite{DBLP:conf/haskell/MistaRH18}---specially the recursive
fields---as well as the individual probability of generating each constructor.
%
For instance, the average number of |Text| data constructors that we will
generate provided that we generated a |Join| data constructor on the previous
level is calculated as follows:
%
$$ M_{Join, Text} = 2\cdot p_{Text} $$
%
where $2$ is the number of holes present when generating a partial ADT value
|Join| (i.e., |Join hole hole|) and $p_{Text}$ is the probability of
individually generating the constructor |Text|.
%
This pattern can be used to build the rest of the mean matrix as well.

%\todo[inline]{Complete this!}

% \begin{align*}
%   E \left[ P_n \right]
%   = \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
% \end{align*}

In this work, we show how to naturally fit structural information beyond ADT
definitions into the prediction mechanism of branching processes.
%
It is our realization that it suffices to simply consider \emph{each different
  pattern matching and functional call as a kind of individual on its own}.
%
In that manner, we can extend our mean matrix $M$ by adding a row and a column
for each different pattern matching and function call as follows:

\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  \hspace{20pt}
  M = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

To extend $M$ to consider the new possible constructions, we need to calculate
$M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
and $j$, respectively.
%
Fortunately, it is possible to follow the same pattern as we did on our original
work, this time considering the amount of recursive pattern variables if it the
case of a pattern matching value being a father individual, and the amount of
recursive arguments if we deal with a abstract interface function call as a
father individual.
%
For instance, the average amount of function calls to |div| that we will produce
provided that we generated a value matching the second pattern of |simplify| as a
father results $2\cdot p_{div}$.
%
Similarly, the average amount of values satisfying the first pattern matching of
|simplify| provided that we generated a |Tag| constructor as a father results
$1 \cdot p_{simplify\# 1}$.

Furthermore, we found that the prediction process can be factored in terms of
two vectors $\beta$ and $\mathcal{P}$, such that $\beta$ represents the number
of recursive sub-terms of each construction that we generate, whereas
$\mathcal{P}$ simply represents the probability of generating that construction
(this probability can be easily calculated based on the generation frequencies
used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}
%
The vector $\beta$ can be obtained by analyzing the shape of each source of
structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that can be calculated separately):

\begin{equation*}
  \hspace{20pt}
  \beta = \qquad \qquad \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      0 \\ % Text
      0 \\ % Sing
      1 \\ % Tag
      2 \\ % Join
      0 \\ % simplify#1
      2 \\ % simplify#1
      0 \\ % hr
      1 \\ % div
      1 \\ % bold
    };
    \draw[densely dashed] ([xshift=-3pt]v-4-1.south west) -- ([xshift=3pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
  \end{tikzpicture}
  \hspace{20pt}
  \mathcal{P} =
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      p_{Text} \\
      p_{Single} \\
      p_{Tag}  \\
      p_{Join} \\
      p_{simplify\#1} \\
      p_{simplify\#2} \\
      p_{hr} \\
      p_{div} \\
      p_{bold} \\
    };
    \draw[densely dashed] ([xshift=-15pt]v-4-1.south west) -- ([xshift=15pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
  \end{tikzpicture}
\end{equation*}

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized.
%
In this light, \dragenp follows the same approach as \dragen, using our adapted
prediction mechanism to tune the generation probabilities of each each source of
structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
This way she can specify, for instance, a uniform distribution of data
constructors, pattern matching values and function calls or, alternative, a
distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.

\paragraph{End-to-End Prediction}
%
% So far we have shown that our prection model is able to predict the distribution
% of data constructors, composite pattern values and interface function calls that
% our automatically derived random generators will produce on average.

It is possible to provide further ``end-to-end'' guarantees on the generated
values if we restrict the generation process to only bare data constructors and
pattern matching values.
%
In particular, we can stop considering pattern matching values as atomic
constructions and start seeing them as compositions of several data
constructors.
%
Then, is possible to obtain the average ``total'' number of data constructors
that our generators will produce, regardless of if they are generated on their
own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$.

The prediction only needs to add the expected number of constructors that we
will additionally generate within pattern matching values to the prediction of
bare data constructors.
%
For instance, we can calculate the total average number of constructors |Text|
and |Join| that we will generate simply by expanding the average number of
generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
corresponding data constructors:
%
{\small
  \begin{align*}
    E^\downarrow\!\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
    E^\downarrow\!\!\left[ Join \right]
    = E\left[ Join \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
  \end{align*}
}
%
Given that each time we generate a value satisfying the first pattern matching
of the function |simplify| we add two |Text| and one |Join| data constructors to
our random value.
%
The case of the second pattern matching of |simplify| follows analogously.

Note that this expansion cannot be applied if we also generate random values
containing abstract interface function calls, as we cannot predict the output of
an arbitrary function without solving the halting problem in the process.

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End: