### 2019-01-30 paper/abstract.tex
General pass. Paper still needs compressing

\begin{abstract}

  % Status
  Automatic generation of random values described by algebraic data types (ADTs)
  is often a hard task.
  % data types (ADTs), % ---specially when considering ADTs which
  % represent HTML pages, images, programs or other structural rich objects.
  % Situation
  State-of-the-art random testing tools can automatically synthesize random data
  generators based on ADTs definitions.
  %
  % In that manner, generated values comply with the structure described by
  % ADTs---something that proves useful when testing software which expects
  % complex input like web pages, images, or even programs.
  %
  In that manner, generated values comply with the structure described by ADTs,
  something that proves useful when testing software which expects complex
  inputs.
  %
  % Problem
  However, it sometimes becomes necessary to generate structural richer ADTs
  values in order to test deeper software layers.
  %
  % Solution
  % In this light, we propose to leverage static information found in the
  In this work we propose to leverage static information found in the codebase
  as a manner to improve the generation process.
  %
  Namely, our generators are capable to consider how programs branch on input
  data as well as how ADTs values are built via interfaces.
  %
  % Concretely, we propose to consider not only ADTs definitions, but also how
  % programs branch on input data as well as how ADTs values get manipulated via
  % interfaces.
  %
  We implement a tool, responsible to synthezise generators for ADTs values
  while providing compile-time guarantees about their distributions.
  %
  Using compile-time predictions, we provide an heuristic that tries to adjust
  the distribution of generators to what developers might want.
  % \todo[inline,author=Ale]{Case study}
  Preliminary experiments of our approach are very encouraging.
\end{abstract}