### 2019-02-01 paper/finalremarks.tex

\section{Final Remarks}

We extended the standard approach for automatically deriving random generators
in Haskell.
%
Our generators are capable of producing complex and interesting random values by
exploiting static structural information found in the user codebase.
%
Based on the theory of branching processes, we adapt our previous prediction
mechanism to characterize the distribution of random values representing the
different sources of structural information that our generators might produce.
%
This predictions let us optimize the generation parameters in compile time,
resulting in an improved testing performance according to our experiments.

In the future we aim to extract stronger invariants from the user codebase into
our automatically derived generators, as well as evaluating how introducing
random mutations into structurally complex values might improve testing
performance, specially when generating ``almost correct'' random data.