### 2019-04-16 papers/haskell19/5-typelevel.tex
Added file

\section{Type-Level Generation Specifications}
\label{sec:typelevel}

\begin{code}
type family Con (c :: Symbol)             :: * -> *
type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
type family Fun (f :: Symbol)             :: * -> *
\end{code}

\begin{code}
data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
\end{code}

\begin{code}
type family (t :: * -> *) apply (a :: *) :: * -> *  where
  (Some t)      apply  a  = t a
  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
  (f otimes n)  apply  a  = (f apply a) otimes n
  (Term t)      apply  a  = Term (t apply a)
  t             apply  a  = t
\end{code}

\begin{code}
type instance Con "Text"        = Con_Text
type instance Pat "simplify" 1  = Pat_simplify_1
type instance Fun "<+>"         = Fun_cat
\end{code}

\begin{code}
type Con' a    = Term  (Con a)
type Fun' a    = Term  (Fun a)
type Pat' a n  = Term  (Pat a n)
\end{code}

\begin{code}
type Html_Spec
  =      Con'  "Text"
  oplus  Con   "Join"         otimes 2
  oplus  Fun'  "br"           otimes 4
  oplus  Fun   "<+>"          otimes 9
  oplus  Pat   "simplify"  1  otimes 3
  oplus  Pat   "simplify"  2  otimes 5
\end{code}

\begin{code}
instance Arbitrary Html where
  arbitrary = sized (genEval at Html_Spec)
\end{code}