### 2019-04-24 papers/haskell19/4-generators.tex
WIP Section 4

\section{Generating Random Constructions}
\label{sec:generators}

% --------------------------------------
% Intro
%
So far we have seen how to encode different constructions representing
interesting values from our target data types.
%
Such representations follow a modular approach, where each different
construction is considered independent from the rest.
%
This modularity allows us to derive each different construction independently,
as well to specify the shape of our target data with with a fine granularity.

In this section we introduce the machinery required to randomly generate the
values encoded using our representations.
%
For this purpose, this step also follows the modular, fine-grained fashion,
resulting in a generation process entirely compositional.
%
In this light, our random generators are built from smaller ones (each one
representing a single construction), and they are based solely on the types we
use to represent the shape of our random data.

Ideally, our aim is to be able to obtain random generators with a behavior
similar to the one presented for |Html| in Section \ref{sec:sources}.
%
If we take a closer look at its definition, we can observe three factors
happening simultaneously:
%
\begin{itemize}
\item We use the QuickCheck's generation size to limit the depth of the
  generated values, reducing it by one on each recursive call of the local
  auxiliary function |gen|.
\item We differentiate between \emph{terminal} and \emph{non-terminal
    constructors}, limiting our generator to pick only among terminal ones when
  we have reached the maximum depth (case |gen 0|).
\item We generate different constructions in a (possibly) different frequency
  that is fixed over time.
\end{itemize}
%
Then, in order to achieve the desired generation behavior, we will need to model
these previous aspects in our modular framework, and in such a way that we do
not compromise the compositionality obtained so far.

% --------------------------------------
% Depth-bounded generators
%
\subsection{Depth-Bounded Modular Generators}

The first obstacle that arises when trying to generate random values with a
limited depth using our approach is related to modularity.
%
If we recall the random generator for |Html| from Section \ref{sec:sources}, we
can observe that the depth parameter |d| is threaded to the different recursive
call of our generator, but always within the scope of the |gen| local function.
%
Since each construction will have an specialized random generator, we cannot
group them as we did before using an internal |gen| function.
%
Instead, we will define a depth-bounded generator data type, wrapping
QuickCheck's |Gen| data type with a external parameter representing the current
maximum recursive depth:

\begin{code}
type BGen a = Int -> Gen a
\end{code}

A |BGen| is essentially a normal |Gen| with the maximum recursive depth as an
input parameter.
%
Using this previous definition, we can generalize QuickCheck's |Arbitrary| class
to work with depth-bounded generators simply as follows:

\begin{code}
class BArbitrary (a :: *) where
  bounded :: BGen a
\end{code}

From now on, we will use this type class as a more flexibe substitute of
|Arbitrary|, given that now we have two parameters to tune: the maximum
recursive depth, and the QuickCheck generation size.
%
Where the former is useful to tune the overall size of our random data, whereas
the latter can be use for tuning the values of the ``leaves'', such as the
maximum length of random |String| or the maximum/minumum random |Int|s.

Here we want to actively remark that, even though we could have reused the
QuickCheck generation size parameter to simultaneously model the maximum
recursive depth and the maximum size of the leaves, doing so would imply
generating random leaves with a decresing size as we move deeper within a random
value, obtaining for instance, random trees with all zeroes on its leaves, or
random lists showing a hight tendency to be ordered in decreasing order.
%
In addition, one can always obtain a trivial |Arbitrary| instance from a
|BArbitrary| one by using QuickCheck's generation size both for the maximum
recursive depth as well as for generating leaves:

\begin{code}
instance BArbitrary a => Arbitrary a where
  arbitrary = sized bounded
\end{code} %$

Even though we have extended QuickCheck's generators to be depth-aware, here we
also need to consider the parametric nature of our representations.
%
In the previous section, we defined each construction representation as being
parametric on the type of the recursive sub-terms, as a manner to defer this
choice until we have specified the overal shape of our target data.
%
Hence, each construction representation is of kind |* -> *|.
%
If we want to define our generators in a modular way, we also need to
parametrize somehow the generation of the recursive subterms!
%
Fortunately, QuickCheck already defines a lifted type class |Arbitrary1| for
types for parametric types of kind |*->*|:

\begin{code}
class Arbitrary1 (f :: * -> *) where
  liftArbitrary :: Gen a -> Gen (f a)
\end{code}

This type class allows us to write modular generators, leaving unspecified the
specific generator used to generate values of its type argument.
%
For instance, the default parametric generator for the data type |Maybe| looks
like:

\begin{code}
instance Arbitrary1 Maybe where
  liftArbitrary gen = frequency
    [  (1,  pure Nothing),  (3,  Just <$> gen) ]
\end{code} %$

Then, we can use the same mechanism for our modular generators, extending it to
be depth-aware as follows:

\begin{code}
class BArbitrary1 (f :: * -> *) where
  liftBGen :: BGen a -> BGen (f a)
\end{code}

Note the similarity between the |Arbitrary1| and |BArbitrary1| type classes.
%
We will use this type class to implement random generators for each construction
we are considering.
%
Recalling our |Html| example, we can finally define modular random generator for
the constructions representing data constructors as follows:

\begin{code}
instance BArbitrary1 Con_Text where
  liftBGen gen d = Con_Text <$> arbitrary

instance BArbitrary1 Con_Sing where
  liftBGen gen d = Con_Sing <$> arbitrary

instance BArbitrary1 Con_Tag where
  liftBGen gen d = Con_Tag <$> arbitrary <*> gen (d - 1)

instance BArbitrary1 Con_Join where
  liftBGen gen d = Con_Join <$> gen (d - 1) <*> gen (d - 1)
\end{code} %$

Note in the previous definitions how each instance is defined to be independent
of the maximum depth (paramtrized by the input variable |d|) and the random
generator used for the recursive sub-terms (parametrized by the input variable
|gen|).
%
Every other non-recursive sub-term can be generated using a normal |Arbitrary|
instance---we use this to generate random |String|s in the previous definitions.

The rest of our construction representations can be generated using the same
idea.
%
For example, the |BArbitrary1| instances for |Fun_bold| and |Pat_simplify_2| are
as follows:

\begin{code}
instance BArbitrary1 Fun_bold where
  liftBGen gen d = Fun_bold <$> gen (d - 1)

instance BArbitrary1 Pat_simplify_2 where
  liftBGen gen d = Pat_simplify_2  <$> arbitrary
                                   <*> gen (d - 1)
                                   <*> gen (d - 1)
\end{code} %$

% \begin{code}
% instance BArbitrary1 Fun_br where
%   liftBGen gen d = pure Fun_br

% instance BArbitrary1 Fun_bold where
%   liftBGen gen d = Fun_bold <$> gen (d - 1)

% instance BArbitrary1 Fun_cat where
%   liftBGen gen d = Fun_cat <$> gen (d - 1) <*> gen (d - 1)

% instance BArbitrary1 Fun_list where
%   liftBGen gen d = Fun_list <$> liftArbitrary (gen (d - 1))
% \end{code} %$

% \begin{code}
% instance BArbitrary1 Pat_simplify_1 where
%   liftBGen gen d
%     = Pat_simplify_1 <$> arbitrary <*> arbitrary

% instance BArbitrary1 Pat_simplify_2 where
%   liftBGen gen d
%     = Pat_simplify_2 <$> gen (d - 1) <*> gen (d - 1)
% \end{code}

Having all the modular generators for each construction in place, we can finally
obtain a concrete depth-aware generator (of kind |*|) for any final
representation |f| as follows:

\begin{code}
instance BArbitrary1 f => BArbitrary (Fix f) where
  bounded d = Fix <$> liftBGen bounded d
\end{code} %$

The previous instance uses the |BArbitrary1| of our final representation to
generate sub-terms recursively using itself (|bounded d|) as the parametrized
input generator, wrapping each recursive sub-term with a |Fix| data constructor.

The machinery developed so far lets us generate single random constructions in a
modular fashion.
%
However, we still need to develop our generation mechanism a bit further in
order to generate composite representations built using the |oplus''| type-level
operator.
%
This is the objective of the next sub-section.

% --------------------------------------
% Encoding Generation Instructions Using Types
%
\subsection{Encoding Generation Instructions Using Types}

\begin{code}
newtype ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
\end{code}

\begin{code}
newtype Term (f :: * -> *) a = Term (f a)
\end{code}

\begin{code}
instance Algebra f a => Algebra (f otimes n) a where
  alg (Freq f) = alg f
\end{code}

\begin{code}
instance Algebra f a => Algebra (Term f) a where
  alg (Term f) = alg f
\end{code}

\begin{code}
instance BArbitrary1 f => BArbitrary1 (f otimes n) where
  liftBGen gen d = Freq <$> liftBGen gen d

instance BArbitrary1 f => BArbitrary1 (Term f) where
  liftBGen gen d = Term <$> liftBGen gen d
\end{code}

\begin{code}
type family FreqOf (f :: * -> *) :: Nat where
  FreqOf (f oplus g)   = FreqOf f + FreqOf g
  FreqOf (f otimes n)  = n * FreqOf f
  FreqOf (Term f)      = FreqOf f
  FreqOf _             = 1
\end{code}

\begin{code}
type family FreqOf' (f :: * -> *) :: Nat where
  FreqOf' (f oplus g)   = FreqOf' f  +  FreqOf' g
  FreqOf' (f otimes n)  = n  *  FreqOf' f
  FreqOf' (Term f)      = FreqOf f
  FreqOf' _             = 0
\end{code}

\begin{code}
instance (BArbitrary1 f, BArbitrary1 g)
  => BArbitrary1 (f oplus g) where
  liftBGen gen d =
    if d > 0
    then frequency
      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen d)
      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen d) ]
    else frequency
      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen d)
      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen d) ]
\end{code}

\begin{code}
freqVal :: forall n dot KnownNat n => Int
\end{code}

\begin{code}
genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)
genFix d = bounded at(Fix f) d
\end{code}

\begin{code}
genEval :: forall f a dot (BArbitrary1 f, Algebra f a) => BGen a
genEval d = eval <$> genFix at f d
\end{code} %$