### 2019-04-29 papers/haskell19/6-casestudies.tex
WIP case studies

\section{Case Studies}
\label{sec:casestudies}

The random generation framework presented throughout this work allows us to
write extensible generators in a very concise way.
%
However, this expresiveness comes attached to a perceptible runtime overhead.
%
In this section we evaluate the implicit cost of composing generators using
three real-world case studies.

As we have shown in Section \ref{sec:generators}, the random generation
process we propose in this paper can be seen as having two phases.
%
First, generate random values from the representation types used to specify the
shape of our data; and then we use their algebras to translate them to the
corresponding values of our target data types.
%
In particular, the tranformation step is expected to pattern match repeatedly
against the |InL| and |InR| constructors of the |oplus| operators when
traversing each construction injection.
%
Because of this, in general we expect a performance impact with respect to
manually-written concrete generators.

As recently analyzed by \citeauthor{KiriyamaOptimizingDTC}, this slowdown is
expected to be linear in the depth of our representation type
\cite{KiriyamaOptimizingDTC}.
%
Hence, one can drastically reduce the runtime overhead by associating each
|oplus| operator in a balanced fashion.
%
So, for instance, instead of writing |(f oplus'' g oplus'' h oplus'' i)|, which
is implicitly parsed as |(f oplus'' (g oplus'' (h oplus'' i)))|; we can
associate constructions as |((f oplus'' g) oplus'' (h oplus'' i))|, thus
reducing the depth of our representation from four to three levels and, in
general, from a $\mathcal{O}(n)$ to a $\mathcal{O}(log(n))$ complexity in the
runtime overhead, where $n$ is the amount of constructions under consideration.
%
To apply this optimization automatically, and relying on more sophisticated
techniques like prefix-free encodings to obtain optimal representations, is a
undergoing path of our future work.

To evaluate our approach, we analyzed the performance of generating random
values using three real-world case studies:
\begin{inparaenum}[(i)]
\item Red-Black Trees, inspired by Okasaki's formulation,
\item Lisp S-expressions, inspired by the package \emph{hs-zuramaru}, and
\item HTML expressions, inspired by the \emph{html} package, and following the
  same structure as our motivating |Html| example.
\end{inparaenum}
%
The expanse of each case study can be summarized as follows:

\begin{table}[H]
  \begin{tabular}{||c||c||c||c||c||}
    \hline
    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
    \hline
    RBT        & 2      & 6      & 5      & 13                  \\ \hline
    SExp       & 6      & 9      & -      & 15                  \\ \hline
    Html       & 4      & -      & 132    & 136                 \\ \hline
  \end{tabular}
\end{table}

For each case study, we benchmarked the execution time required to generate 1000
random values using both manually-written concrete generators, and modular
generators obtained using our approach.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,5}
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, Html}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
      \legend{Manual, Modular/Linear, Modular/Balanced}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}