### 2019-05-01 papers/haskell19/2-sources.tex
Section 2

\section{Random Testing with Source Hints}
\label{sec:sources}

% --------------------------------------
% Motivating Example
%
This section introduces the common workflow for generating random values using
automatically derived QuickCheck generators, along with its common limitations.
%
To illustrate this, let us consider the following data type definition encoding
basic HTML pages:

\begin{code}
data Html
  =  Text  String
  |  Sing  String
  |  Tag   String Html
  |  Join  Html Html
\end{code}

This type allows us to build HTML pages via four possible data constructors:
|Text| is used for plain text values; |Sing| and |Tag| represent singular and
paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
one after another.
%
Note that the constructors |Tag| and |Join| are recursive, as they have at least
one field of their own type.
%
In this paper, and in the scope of random generation in general, we will refer
to any non-recursive data constructor as \emph{terminal}, and
\emph{non-terminal} in any other case.

Then, using this representation the example page\linebreak
%
\texttt{<html>hello<hr><b>world!</b></html>}
%
can be encoded simply as:
%
\begin{code}
Tag "html" $  Text "hello"
              iJoin Single "hr"
              iJoin Tag "b" $ Text "world!"
\end{code} %$

% --------------------------------------
% Random Generators
%
Then, in order to generate random values of type |Html|, QuickCheck requires us
to provide a random data generator for it.
%
For this purpose, QuickCheck defines the |Arbitrary| type-class for default
defining random generators as:

\begin{code}
class Arbitrary (a :: *) where
  arbitrary :: Gen a
\end{code}
%
where the overloaded symbol |arbitrary :: Gen a| denotes a monadic generator for
values of type |a|.
%
Using this mechanism, we can define a sensible random generator for our |Html|
data type as follows:

\begin{code}
instance Arbitrary Html where
  arbitrary = sized gen
    where
      gen 0 = frequency
        [  (2,  Text    <$> arbitrary)
        ,  (1,  Sing    <$> arbitrary) ]
      gen d = frequency
        [  (2,  Text    <$> arbitrary)
        ,  (1,  Sing    <$> arbitrary)
        ,  (4,  Tag     <$> arbitrary  <*> gen (d-1))
        ,  (3,  Join    <$> gen (d-1)  <*> gen (d-1)) ]
\end{code} %$

At the top level, this definition parametrizes the generation process using
QuickCheck's |sized| combinator, which lets us build our generator using an
auxliary local function |gen :: Int -> Gen Html|.
%
The |Int| passed to |gen| is known as the \emph{generation size}, and is
threaded seamlessly by QuickCheck on each call to arbitrary.
%
Then, we use this parameter to limit the maximum amount of recursive calls that
our generator can perform, and thus the maximum depth of the generated values.
%
If the generation size is positive (case |gen d|), our random generator picks a
random constructor from our |Html| data type with a given generation frequency
(using QuickCheck's |frequency| combinator), and proceeds to fill its fields
using randomly generated sub-terms---here using Haskell's applicative notation
and the default |Arbitrary| instance for |String|s.
%
For the case of the recursive sub-terms, this generator simply calls itself
recursively with a smaller generation size (|gen (d-1)|).
%
This process repeats until we reach the base case (|gen 0|) on each recursive
sub-term.
%
At this point, our random generator is limited to only pick among the |Html|
terminal constructors, hence terminating the generation process.

% --------------------------------------
% Automatic derivation tools
%
As one can observe, the previous definition is quite mechanical, and depends
only on the generation frequencies we choose for each constructor.
%
Fortunately, there exists a variety of meta-programming tools for obtaining
random generators automatically based on their types.
%
Tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random generators
using the same procedure as the one shown above.
%
\emph{MegaDeTH} simply picks among data constructors uniformingly, whereas
\emph{DRAGEN} analyzes the type definitions and tunes the generation frequencies
based on a simulation-based optimization approach, in order to obtain random
distributions of values based on the users' desires.

% --------------------------------------
% Limitation
%
These tools provide a useful mechanism for automating the boring and error-prone
task of writing random generators by hand.
%
However, this simple derivation procedure which only considers the structural
information encoded on data type definitions, and generates one single
constructor at a time; is often too generic to synthesize useful generators in
some practical scenarios.
%
For instance, because of the types we use for deriving our random generators are
too lax or, alternatively, because our properties need to be tested against very
specific patterns of input cases in order to trigger bugs.

This problem, however, can be alleviated by considering additional sources of
structural information from our codebase other than just data constructors.
%
We proceed to illustrate this idea for consideting the structural information
found on abstract interfaces and functions' pattern matchings.

% --------------------------------------
% Abstract interfaces
%
\paragraph{Extracting Information From Interfaces}

A common choice when implementing abstract data types is to transfer the
resposiblity of preserving their invariants to the functions on their abstract
interface.
%
Take for example our |Html| data type.
%
Instead of defining a different data constructor for each possible HTML tag, we
opted for a very small generic representation that can be extended via a set of
combinators making use of this generic interface:

\begin{code}
br :: Html
br = Sing "br"

bold :: Html -> Html
bold = Tag "b"

list :: [Html] -> Html
list []  = Text "empty list"
list xs  = Tag "ul" $ foldl1 Join (Tag "li" <$> xs)

(<+>) :: Html -> Html -> Html
(<+>) x y = Join x (Join br y)
\end{code} %$

These functions let us create interesting structured values representing HTML
constructions like linebreaks (|br|), bold blocks (|bold|), unordered lists
(|list|) and concatenation of values one below another (|<+>|).

Note how difficult would be to generate random values containing, for example,
structurally valid HTML lists if we only consider the structural information
encoded in our |Html| data type definion.
%
After all, much of the valid structure of HTML has been encoded on its abstract
interface.

An automatically derived random generator could integrate this structural
information into the generation process by simply creating random values arising
from applying such functions using randomly generated inputs:

\begin{code}
  gen d = frequency
    [ ...
    , (1, pure br)
    , (5, bold   <$> gen (d-1))
    , (2, list   <$> listOf (gen (d-1)))
    , (3, (<+>)  <$> gen (d-1) <*> gen (d-1)) ]
\end{code}%$

where the symbol (|...|) represents the rest of the possible sub-generators for
each data constructor seen before.
%
From now on, we will refer to each possible choice given to |frequency| as a
different \emph{random construction}, as we are not considering only generating
only single data constructors anymore.

% --------------------------------------
% Functions' pattern matchings
%
\paragraph{Extracting Information From Patterns}

A different challenge appears when we try to test functions comprising complex
pattern matchings. Consider for instance the following function for simplifying
|Html| values:

\begin{code}
simplify :: Html -> Html
simplify (Text t1 iJoin Text t2) = Text (t1 ++ t2)
simplify ((Text t1 iJoin x) iJoin y)
  = simplify (Text t1 iJoin simplify (x iJoin y))
simplify (x iJoin y) = simplify x iJoin simplify y
simplify (Tag t x) = Tag t (simplify x)
simplify x = x
\end{code}

This function simply traverses |Html| values, joining together every contiguous
pair of |Text| constructors.
%
Ideally, we would like to put approximately the same testing effort info each
clause of |simplify|, or perhaps even more to first two ones, since those are
the ones performing actual simplifications.
%
However, these two clauses are the most difficult ones to test in practice.
%
The probabilty of generating a random value satisfying nested patterns decreases
multiplicatively with the number of constructors we match against to
simultaneously.
%
In our tests, we were not able to exercise any of these two patterns more than
6\% of the overall testing time (and considering recursive sub-calls), using
random generators derived using either \emph{MegaDeTH} and \emph{DRAGEN}.
%
As expected, most of the random test cases were exercising the simplest (an
rather uninteresting) patterns of this function.

To solve this issue, we could opt to consider each complex pattern as a new kind
of random construction.
%
In this light, we can simply generate values satisfying pattern directly by
returning its corresponding expression, where each variable or wildcard pattern
is filled using a random sub-expression:

\begin{code}
  gen d = frequency
    [ ...
    , (2, do  t1  <- arbitrary
              t2  <- arbitrary
              return (Text t1 iJoin Text t2))
    , (4, do  t2  <- arbitrary
              x   <- go (d-1)
              y   <- go (d-1)
              return ((Text t1 iJoin x) iJoin y))]
\end{code}%$

% --------------------------------------
% Wrapping up
%
In general, we think the solution we is far from ideal, as in many cases we
would like to select which random constructions are of our interest when testing
different properties or sub-systems.
%
In this light, having ``monolithic'' random generators derived automatically
would complicate our desired workflow.
%
We would need to derive different concrete random generators for each use case
instead of having a single |Arbitrary| instance that can be used everywhere.
%
In contrast, in this work we opted for developing an extensible fashion, where
the generation machinery is automatically derived only once, and every variant
of our random generators is expressed at the type-level.
%
The following sections of this paper cover our alternative approach in depth.