### 2019-05-01 papers/haskell19/8-conclusions.tex
WIP conclusions

\section{Conclusions}
\label{sec:conclusions}

We presented a novel approach for automatically deriving flexibe random
generators inspired by the seminal work on Data Types \`a la Carte.
%
Furthermore, we incorporate valuable structural information into our generation
process by considering not only data constructors but also the information
statically available in abstract interfaces and functions' pattern matchings.

In the future, something something...