### 2019-05-04 papers/haskell19/8-conclusions.tex

\section{Conclusions}
\label{sec:conclusions}

We presented a novel approach for automatically deriving flexibe random
generators inspired by the seminal work on Data Types \`a la Carte.
%
Furthermore, we incorporate valuable structural information into our generation
process by considering not only data constructors but also the information
statically available in abstract interfaces and functions' pattern matchings.

In the future, we aim to minimize the runtime overhead of our technique, using
perhaps prefix-free encodings based on generation frequencies, as a way to
reduce the representation bureacracy as much as possible.
%
Additionally, we are interested on including additional sources of structural
infrmation from users' codebases, e.g., locally defined patterns or sequencies
of functions applications.
%
Interesting thoughts to motivate our future reseach.