### 2019-05-04 papers/haskell19/abstract.tex

Generating ``good'' random values described by algebraic data types is often
quite intricate.
%
State-of-the-art tools for synthesizing random generators serve the valuable
purpose of helping with this task, while providing different level of invariants
imposed over the generated values.
%
However, none of them are built neither for composability nor extendability, a
useful feature when the shape of our random data makes a difference while
testing different properties or sub-systems.
%
% Instead, users are forced to decide wether to manually write their random
% generators or to use rigid automatically derived ones.

In this work, we develop an extensible framework for deriving compositional
generators, which can be easily combined in diferent ways in order to fit the
developers' demands, using a simple type-level domain-specific language.
%
Our framework is based on familiar ideas from the \`a la Carte technique for
writing composable interpreters in Haskell, and extended with the required
machinery for it to work in the scope of random generation of values.
%
Additonally, we provide an implementation of our ideas, and evaluate its
performance using real-world examples.