### 2019-05-07 papers/haskell19/abstract.tex

Generating ``good'' random values described by algebraic data types is often
quite intricate.
%
State-of-the-art tools for synthesizing random generators serve the valuable
purpose of helping with this task, while providing different level of invariants
imposed over the generated values.
%
However, none of them are built neither for composability nor extendability, a
useful feature when the shape of our random data makes a difference while
testing different properties or sub-systems.
%
% Instead, users are forced to decide wether to manually write their random
% generators or to use rigid automatically derived ones.

In this work, we develop an extensible framework for deriving compositional
generators, which can be easily combined in different ways in order to fit the
developers' demands, using a simple type-level description language.
%
Our framework relies on familiar ideas from the \`a la Carte technique for
writing composable interpreters in Haskell.
%
In this light, we adapt this technique with the machinery required in the scope
of random generation.
%
We provide an implementation of our ideas, and evaluate its performance using
real-world examples.
%