### 2019-05-11 papers/haskell19/8-conclusions.tex
Last minute changes

\section{Conclusions}
\label{sec:conclusions}

We presented a novel approach for automatically deriving flexible random
generators inspired by the seminal work on Data Types \`a la Carte.
%
Furthermore, we incorporate valuable structural information into our generation
process by considering not only data constructors, but also the structural
information statically available in abstract interfaces and functions' pattern
matchings.
%

In the future, we aim to minimize the runtime overhead of our technique, using
perhaps prefix-free encodings based on generation frequencies, as a way to
reduce the representation bureaucracy as much as possible.
%
In addition, using type-level representations for each random construction
leaves the door open for extending our approach with a composable mechanism for
predicting the distributions of random values.
%
This would be possible by overloading each representation type with stochastic
information describing its branching behavior---interesting thoughts to motivate
our future research.

% Additionally, we are interested on including additional sources of structural
% information from the users' codebases, e.g., locally defined patterns or
% sequences of functions applications